### 2021-06-08 memo.tex
better treatment of monopoles of so(2nc)

So we would like to determine that in the $SO$ SQCD,
but the direct analysis of dynamical monopoles in this theory is hard.
For this purpose, we make the following deformations which do not change the symmetry structure:
\begin{itemize}
\item We reduce the flavor symmetry from $su(2n_f)$ to $usp(2n_f)$.
\item We add an adjoint scalar $\Phi_{[ab]}$ and the interaction $\psi^{ai}_\alpha \psi^{bj}_\beta J_{ij} \Phi_{ab}\epsilon^{\alpha\beta} +cc$, where $J_{[ij]}$ is the constant matrix for the $usp(2n_f)$ part.
\item We give a vev to $\Phi_{ab}$ to break $so(2n_c)$ to $so(2)^{n_c}$.
\end{itemize}
The 't Hooft line of the original $SO(2n_c)$ theory is in the `vector' class
$(1,0,\ldots,0)$ under $so(2)^{n_c}$,
and therefore the dynamical monopole we need to analyze has the charge $(2,0,\ldots,0)$.

Since the fermions are in the vector representation of $so(2n_c)$,
they are charge 2 under the $so(2)$, and should give rise to fermionic zero modes transforming in \begin{equation}
V_{2} \otimes R_{2n_f} 
\end{equation} where $V_2$ is the doublet of spacetime $so(3)$, 
$R_{2n_f}$ is the fundamental of $usp(2n_f)$,
and we need to impose the reality condition using the pseudoreality of both factors,
so that there are $4n_f$ Majorana fermion in total.

I haven't analyzed the general case; let's take $n_f=1$ and $n_f=2$ as examples.
When $n_f=1$, there are $4$ Majorana fermions.
Quantizing them, we find the monopoles in \begin{equation}
V_2 \otimes \mathbf{1} \oplus \mathbf{1}\otimes R_2.
\end{equation}
I think we should take the spacetime scalar as indicative, (which needs to be justified)
so it has the `vector' charge under $usp(2)$ flavor symmetry.
Therefore the 1-form symmetry is extended.

When $n_f=2$, there are $8$ Majorana fermions, 
and spatial $su(2)$ and flavor $usp(4)$ are embedded into $so(8)$ via \begin{equation}
su(2)\times usp(4) \simeq so(3)\times so(5) \subset so(8).
\end{equation}
Therefore, the monopoles are easily seen to be in \begin{equation}
V_3\otimes \mathbf{1} \oplus V_2\otimes R_2 \oplus V_1\otimes R_5.
\end{equation}
Again taking the spacetime scalar part, it is in the `adjoint' charge under the $usp(4)$ flavor symmetry.
Therefore the 1-form symmetry is not extended.

>>>>

So we would like to determine that in the $SO$ SQCD,
but the direct analysis of dynamical monopoles in this theory is hard.
For this purpose, we make the following deformations which do not change the symmetry structure:
\begin{itemize}
\item We reduce the flavor symmetry from $su(2n_f)$ to $usp(2n_f)$.
\item We add an adjoint scalar $\Phi_{[ab]}$ and the interaction $\psi^{ai}_\alpha \psi^{bj}_\beta J_{ij} \Phi_{ab}\epsilon^{\alpha\beta} +cc$, where $J_{[ij]}$ is the constant matrix for the $usp(2n_f)$ part.
\item We give a vev to $\Phi_{ab}$ to break $so(2n_c)$ to $so(2)^{n_c}$.
\end{itemize}
The 't Hooft line of the original $SO(2n_c)$ theory is in the `vector' class
$(1,0,\ldots,0)$ under $so(2)^{n_c}$,
and therefore the dynamical monopole we need to analyze has the charge in the `adjoint' class.

The monopoles associated to the breaking of $G$ to its Cartan were analyzed in many places, see e.g.~\cite{Gauntlett:1996cw}.\footnote{%
There were many competing papers from the same period concerning the application to Seiberg-Witten;
the analysis of the monopoles itself is much older and I should find proper original references for it too.
}
The upshot is that you can embed the standard spherically-symmetric 't Hooft-Polyakov monopole using any simple root into $G$ and have a monopole solution without additional bosonic moduli.

So we consider a monopole with the charge $(1,1,0,\ldots,0)$.
This uses $su(2)_1$ within $so(4)\simeq su(2)_1\times su(2)_2$ within $so(2n_c)\subset so(4)\times so(4n_c-4)$.

Since the fermions are in the vector representation of $so(2n_c)$,
the charged part under $su(2)_1$ is of charge 1 and in the doublet of $su(2)_2$.
Therefore it gives rise to fermionic zero modes transforming in \begin{equation}
V_{2} \otimes R_{2n_f} 
\end{equation} where $V_2$ is the doublet of $su(2)_2$, 
$R_{2n_f}$ is the fundamental of $usp(2n_f)$,
and we need to impose the reality condition using the pseudoreality of both factors,
so that there are $4n_f$ Majorana fermion in total.

I haven't analyzed the general case; let's take $n_f=1$ and $n_f=2$ as examples.
When $n_f=1$, there are $4$ Majorana fermions.
Quantizing them, we find the monopoles in \begin{equation}
V_2 \otimes \mathbf{1} \oplus \mathbf{1}\otimes R_2.
\end{equation}
%I think we should take the spacetime scalar as indicative, (which needs to be justified)
It has the `vector' charge under $usp(2)$ flavor symmetry or is a doublet under $su(2)_2$.
In either case, the $-1$ in $SO(2n_c)\times USp(2n_f)$ acts nontrivially,
and the 1-form symmetry is extended.

When $n_f=2$, there are $8$ Majorana fermions, 
and  $su(2)_2$ and flavor $usp(4)$ are embedded into $so(8)$ via \begin{equation}
su(2)\times usp(4) \simeq so(3)\times so(5) \subset so(8).
\end{equation}
Therefore, the monopoles are easily seen to be in \begin{equation}
V_3\otimes \mathbf{1} \oplus V_2\otimes R_2 \oplus V_1\otimes R_5.
\end{equation}
Again,
it is in the `adjoint' charge under the $usp(4)$ flavor symmetry.
Therefore the 1-form symmetry is not extended.
